---
title: "Layout"
description: |
  Organize your UI
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

With inputs and outputs established in the UI, we can arrange them with various layout functions.

# Pre-defined Layouts

Some are pre-set like `sidebarLayout()`.

```{r}
ui <- fluidPage(

  title,
  src,
  
  # A sidebarLayout with sidebarPanel and mainPanel
  sidebarLayout(
    sidebarPanel(),
    mainPanel()
    )

)
```

Place content within `sidebarPanel()` and `mainPanel()`.

```{r}
ui <- fluidPage(
  
  title,
  src,
  
  # A sidebarLayout with sidebarPanel and mainPanel
  sidebarLayout(
    sidebarPanel(
      txt_box,
      txt_disp
    ),
    mainPanel(
      tbl_disp
    )
  )
)

```

# Rows and Columns

In lieu of pre-defined layouts, you can build your layout with rows and columns using `fluidRow()` or `fixedRow()`. `columns()` are created within `*Row()` . 

<aside>
Set column widths using an integer between 1 and 12. The maximum number of units within a row is 12.
</aside>

You can also embed rows and columns within pre-defined layouts like in the example below.

```{r}
ui <- fluidPage(
  
  title,
  src,
  
  # A sidebarLayout with sidebarPanel and mainPanel
  sidebarLayout(
    sidebarPanel(
      txt_box,
      txt_disp
    ),
    mainPanel(
      # create columns within fluidRow or fixedRow 
      fluidRow(
        column(
          width = 6,
          h3("Table"),
          tbl_disp
        ),
        column(
          width = 6,
          h3("Plot")
        )
      )
    )
  )
  
)
```





