---
title: "Packages II"
description: |
  Tables with DT
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
```
Start with a new file: `intro-shiny/babynames/04-packages-02/app-baby-00.R`

Allow users to sort or filter your tables without having to expand the UI. Simply use packages such as [DT](https://rstudio.github.io/DT/) or `Reactable` to present data frames as interactive tables. 

# DT

DT stands for data table (not to be confused with the wrangling/processing package `data.table`). Like, `Plotly`, DT is part of the htmlwidgets family. Developers have taken the JavaScript version of DT and bound it with R. We don't have to learn JS in order to experience the benefits of it!

Load package
```{r}
library(DT)
```

Update the output and render functions
```{r}
# UI

title <- titlePanel("U.S. Baby Names")
src <- p("Source:",
         tags$a(href = "https://www.ssa.gov/oact/babynames/limits.html",
                "Social Security Administration"))
txt_box <- textInput("name", 
          label = "Enter name",
          placeholder = "Jane")
txt_disp <- textOutput("name_entered")

# in lieu of tableOutput(), use DT's version
tbl_disp <- DTOutput("main_table")
```

```{r}
# Server

  # in lieu of renderTable(), use DT's version
  output$main_table <- renderDT({
    filtered_df()
  })
```

# Honorable Mention: Reactable





