---
title: "Packages II"
description: |
  Tables with DT
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```
Start with a new file: `intro-shiny/babynames/04-packages-02/app-baby-04.R`

Allow users to sort or filter your tables without having to expand the UI. Simply use packages such as [DT](https://rstudio.github.io/DT/) or [Reactable](https://glin.github.io/reactable/) to present data frames as interactive tables. 

# DT

DT stands for data table (not to be confused with the wrangling/processing package `data.table`). Like, `Plotly`, DT is part of the [htmlwidgets](https://www.htmlwidgets.org/) family. Developers have taken the JavaScript version of DT and bound it with R. We don't have to learn JS in order to experience its benefits!

Load package

```{r}
library(DT)
```

Update the output and render functions

```{r}
# UI

# in lieu of tableOutput(), use DT's version
tbl_disp <- DTOutput("main_table")

```

```{r}
# Server

  # in lieu of renderTable(), use DT's version
  output$main_table <- renderDT({
    filtered_df()
  })
```

## Options

DT offers options to customise the appearance of your table without manipulating the data itself.

```{r}
# Server

  output$main_table <- renderDT({
    df <- filtered_df()
    
    # customize the appearance of the DT
    datatable(df,
              colnames = str_to_title(str_replace_all(colnames(df), "_", " ")),
              options = list(pageLength = 20,
                             lengthMenu = c(20, 50, 100))
              )
    
  })
```

## Format Styling

It's possible to style the table with some helper functions.
https://rstudio.github.io/DT/functions.html
https://rstudio.github.io/DT/010-style.html

```{r}
# Server

 output$main_table <- renderDT({
    df <- filtered_df()
    
    datatable(df,
              colnames = str_to_title(str_replace_all(colnames(df), "_", " ")),
              options = list(pageLength = 20,
                             lengthMenu = c(20, 60, 100))
              ) %>%
      
      # conditional styling of cell value. The color of 'name' is based on 'sex'
      formatStyle('name',
                  'sex',
                  color = styleEqual(unique(df$sex), c('dodgerblue', 'orange'))
      ) %>% 
      
      # add a color bar to a numeric column.
      formatStyle('count',
                  background = styleColorBar(df$count, 'rgba(0, 128, 255, .2)'),
                  backgroundSize = '100% 90%',
                  backgroundRepeat = 'no-repeat',
                  backgroundPosition = 'center'
                  )
    
  })
```


## Extensions

# Honorable Mention: Reactable





