---
title: "Packages I"
description: |
  Plots with ggplot2 and Plotly
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

Shiny provides the basic tools for client-server interaction, but a lot of parts that make up an app--how it wrangles and displays data, can come from packages of your choice. All of the packages introduced in the R Learning Series so far (`dplyr`, `ggplot2`, `plotly`, `data.table`, `R markdown`, `sf`, `leaflet`) can be included in Shiny.

# Plots

`ggplot2` can be used with the standard `plotOutput()` and `renderPlot({})` functions. 

## ggplot2

```{r}
# Server

  output$plot <- renderPlot({
    filtered_df() %>% 
      ggplot(aes(x = year, y = count, color = state)) +
        geom_line() +
        facet_wrap(vars(sex), nrow = 2, scales = "free_y")
  })

```

To add more content to our plot (and table), enable the dropdown menu to allow the user to select multiple states.

```{r}
# UI 

  # add multiple argument to allow more than one selection
  selectInput("state",
              label = 'Select State',
              choices = unique(df$state),
              selected = 'Washington',
              multiple = TRUE)
```

```{r}
# Server

 # allow filtering of multiple states with `%in%`
  filtered_df <- eventReactive(input$go, {
    df %>% 
      filter(name == clean_name() & state %in% input$state)
  })
```

## Plotly

[Plotly](https://plotly.com/ggplot2/), a package that helps create interactive graphs, can be used to convert a static `ggplot2` graph into an interactive one. [Plotly](https://plotly.com/r/) on its own can also be used in lieu of ggplot2. 

To render Plotly graphs in Shiny, use `plotlyOutput()` and `renderPlotly({})` in lieu of the standard plot output and render functions. 

Include the `plotly` package in the global section.

```{r}
library(plotly)
```

Convert existing plot output and render functions to their respective Plotly equivalent.

```{r}
# UI

  # convert plotOutput to a Plotly Output
  plotlyOutput("plot")

```

```{r}
# Server
  
  # convert renderPlot to renderPlotly
  output$plot <- renderPlotly({
    
    # create ggplot object
    p <- filtered_df() %>% 
      ggplot(aes(x = year, y = count, color = state)) +
        geom_line() +
        facet_wrap(vars(sex), nrow = 2, scales = "free_y")
    
    # wrap ggplot object with ggplotly
    ggplotly(p)
  })

```

# Tidy evaluation




