---
title: "Reactivity"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Reactive Programming

Shiny follows a reactive programming paradigm[^1]. We don't need to command Shiny to update itself, rather, it will automatically update outputs if an input changes. More specifically, Shiny will only update the dependencies of that particular input. 

[^1]: https://mastering-shiny.org/basic-reactivity.html#reactive-programming

Shiny maximizes the work not done! In other words, it's lazy and will only do the most minimal work required to update outputs.

## Reactives

In lieu of variables and the typical function that we'd write in a standard R script, R code within the server is often placed inside reactive expressions using `reactive({})`. 

```{r}
server <- function(input, output) {
  
  # add reactive expression creating a filtered data frame
  filtered_df <- reactive({
    df %>% 
      filter(name == input$name & state == input$state)
  })
  
  output$name_entered <- renderText({
    c("You entered:", input$name)
  })
  
  output$main_table <- renderTable({
    # call reactive expression
    filtered_df()
  })
  
  
}
```

Reactives help to:

  - not repeat ourselves
  - break-down complex or lengthy code

Reactives, like standard functions:

  - are assigned to a variable
  - are called upon

but unlike the average function, they monitor `input` and dependencies. If reactive sources change, it will update; otherwise, they are lazy.

```{r}
server <- function(input, output) {
  
  # add reactive expression to munge name before using in filtered_df()
  clean_name <- reactive({
    input$name %>% 
      str_replace_all(" ", "") %>% 
      str_trim() %>% 
      str_to_title()
  })
  
  # call clean_name() in lieu of input$name
  filtered_df <- reactive({
    df %>% 
      filter(name == clean_name() & state == input$state)
  })
  
  output$name_entered <- renderText({
    c("You entered:", input$name)
  })
  
  output$main_table <- renderTable({
    filtered_df()
  })
  
  
}
```

## Observers

In contrast to the lazy `reactive({})`, `observe({})` is eager and also forgetful — they run as soon as they possibly can and they don’t remember their previous action[^2].

[^2]: https://mastering-shiny.org/reactivity-objects.html#observers-details

Unlike reactives, observers:

  - aren't assigned to a variable
  - aren't called on
  - don't return values
  
Instead, observers are used primarily for their side-effects.

### Side-effects

Sometimes we want our app to perform an action that is beyond our app's environment. Examples of side-effects can include: writing and saving a file to disk, copying and pasting files, updating a web API, or updating a database[^3].

[^3]: https://mastering-shiny.org/basic-reactivity.html#observers

# Delaying Response

Specify when a reaction of a reactive or observer should take place with `eventReactive({})` or `observeEvent({})`



