---
title: "Reactivity"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Reactive Programming

Shiny follows a reactive programming paradigm[^1]. We don't need to command Shiny to update itself, rather, it will automatically update outputs if an input changes. More specifically, Shiny will only update the dependencies of that particular input. 

[^1]: https://mastering-shiny.org/basic-reactivity.html#reactive-programming

Shiny maximizes the work not done! In other words, it's lazy and will only do the most minimal work required to update outputs.

## Reactives

In lieu of variables and the typical function that we'd write in a standard R script, R code within the server is often placed inside reactive expressions using `reactive({})`. 

```{r}
server <- function(input, output) {
  
  # add reactive expression creating a filtered data frame
  filtered_df <- reactive({
    df %>% 
      filter(name == input$name & state == input$state)
  })
  
  output$name_entered <- renderText({
    c("You entered:", input$name)
  })
  
  output$main_table <- renderTable({
    # call reactive expression
    filtered_df()
  })
  
  
}
```


## Observers

# Delaying Response





